OpenMPIR

### 2017-09-07 paper.tex
Made cilk backend more obvious

\begin{abstract}
Optimizing compilers for task-level parallelism are still in their infancy.
This work implements a frontend which compiles a subset of OpenMP tasks to
Tapir. This enables analyses and optimizations previously inaccessible to
OpenMP codes. Initial performance results for the Barcelona OpenMP tasking
suite and show performance improvements over existing OpenMP compiler
implementations. 
\end{abstract}

@@

In this work, we put that suggestion to the test, implementing OpenMP tasks by
compiling them to Tapir instructions. This enables us to target non-openmp
runtimes and compare performance to existing OpenMP implementations. Specifically,
we run our prototype implementation on the Barcelona OpenMP task suite
\cite{barcelona}. We discuss the kinds of optimizations this enables, and how
the work could be extended to include other parallelism constructs and semantics
in OpenMP. We also discuss the possibility of extending this work to other
programming models, which would help to alleviate some of the fragmentation
issues discussed above.

@@

\begin{itemize}
\item A prototype frontend from OpenMP task constructs to Tapir IR
\item Benchmarks of the implementation, showing performance improvements over
existing OpenMP implementations
\item An initial investigation into the reasons for the performance improvement
\end{itemize}

Our hope for this paper is that it moves the community towards a shared IR that
can be used for many programming models. This work presented in this paper only
represents a small step in that direction, and there is still a herculean amount
of work required. We argue that the benefits vastly outweigh the costs, and
hope to see more work follow suite.

@@

In this section we discuss a significant challenge in HPC: fragmentation of
parallel programming models. There are many, many, many ways to write
parallel programs \cite{....}. From libraries, to language extensions, to
standalone languages, to combinations of the above, fragmentation of parallel
programming models is a problem that is only getting worse. This means that
when writing tooling, optimizations, or analyses, one has to choose which model
to target. One common difficulty among all of these is that reasoning about
parallel programs is hard \cite{}, and compilers can help \cite{}. Therefore
any duplicated work is extremely costly.  This is exactly analogous the problem
LLVM helps to solve.  Many optimizations and analyses can be applied across a
wide range of programming languages and hardware, but until there was an
effective IR, those were often duplicated, resulting in a lot of lost
opportunity.

@@

Specifically, Tapir introduces the \texttt{detach}, \texttt{reattach}, and
\texttt{sync} instructions. \wmnote{How in detail should I go here? I can talk
about these instructions on a broad level as a target for parallel tasks, or
more concretely define what they do. However, we don't seem discuss CFG's
before this so I'm not sure what level of introduction you're thinking of.}
\gsnote{I was thinking roughly a page should be enough. Maybe a summary of what
Tapir is followed by an example of an optimization it enables. I like both the
CSE and tail-call optimization examples from your PPoPP paper. Feel free to 
define CFG and any other terms you need to.}

@@

>>>>

\begin{abstract}
Optimizing compilers for task-level parallelism are still in their infancy.
This work implements a frontend which compiles a subset of OpenMP tasks to
Tapir, an extension to LLVM that represents fork-join parallelism. This enables
analyses and optimizations previously inaccessible to OpenMP codes, as well as
the ability to target other runtimes. Using a Cilk runtime backend, we compare
to existing OpenMP implementations. Initial performance results for the
Barcelona OpenMP tasking suite and show performance improvements over existing
OpenMP compiler implementations. 
\end{abstract}

@@

In this work, we put that suggestion to the test, implementing OpenMP tasks by
compiling them to Tapir instructions. This enables us to target non-openmp
runtimes and compare performance to existing OpenMP implementations. For this
paper, our prototype implementation uses a Cilk runtime as a backend, and we run
our prototype implementation on the Barcelona OpenMP task suite
\cite{barcelona}. We discuss the kinds of optimizations this enables, and how
the work could be extended to include other parallelism constructs and semantics
in OpenMP. We also discuss the possibility of extending this work to other
programming models, which would help to alleviate some of the fragmentation
issues discussed above.

@@

\begin{itemize}
\item A prototype frontend from OpenMP task constructs to Tapir IR
\item Benchmarks of the implementation using a Cilk runtime backend, showing
performance improvements over existing OpenMP implementations
\item An initial investigation into the reasons for the performance improvement
\end{itemize}

Our hope for this paper is that it moves the community towards a shared IR that
can be used for many programming models. This work presented in this paper only
represents a small step in that direction, and there is still a herculean amount
of work required. We argue that the benefits vastly outweigh the costs, and
hope to see more work follow suit.

@@

In this section we discuss a significant challenge in HPC: fragmentation of
parallel programming models. There are many, many, many ways to write
parallel programs \cite{....}. From libraries, to language extensions, to
standalone languages, to combinations of the above, fragmentation of parallel
programming models is a problem that is only getting worse. This means that
when writing tooling, optimizations, or analyses, one has to choose which model
to target. One common difficulty among all of these is that reasoning about
parallel programs is hard \cite{}, and compilers can help \cite{}. This is
exactly analogous the problem LLVM solves: compiler optimizations are
hard to do well, so sharing them is valuable. Tapir is a natural extension

With fragmentation comes questions of compatibility, or composability, of 
different programming models. To illustrate this issue, consider the example of
a program using a library that uses OpenMP to implement parallelism internally. 
If said program uses a different programming model for parallelism, e.g. Cilk
, there is currently no good solution for

@@

Specifically, Tapir introduces the \texttt{detach}, \texttt{reattach}, and
\texttt{sync} instructions. \wmnote{How in detail should I go here? I can talk
about these instructions on a broad level as a target for parallel tasks, or
more concretely define what they do. However, we don't seem discuss CFG's
before this so I'm not sure what level of introduction you're thinking of.}
\gsnote{I was thinking roughly a page should be enough. Maybe a summary of what
Tapir is followed by an example of an optimization it enables. I like both the
CSE and tail-call optimization examples from your PPoPP paper. Feel free to 
define CFG and any other terms you need to. If that's too much to ask on such 
short notice just let me know and I'll help out.}

@@

While OpenMP and Cilk have similar fork-join style semantics that Tapir
implements naturally, many allow a more flexible mechanism for synchronization:
futures \cite{qthreads, hpx, chapel, charm++}. The primary difference
between the fork-join parallelism represented by Tapir and these approaches
is that \emph{child tasks can outlive their parents}. Figuring out how
to represent this class of parallelism in Tapir represents a significant
challenge for future work.
